---
title: "Mapping free school meals in York, UK. 2021"
output: html_notebook
---



# Task

What can we learn about child poverty in York from published free school meals data?

# Data Sources

Data Source: https://data.yorkopendata.org/dataset/free-school-meals/resource/03b0ae33-f6fe-4431-94f8-4b811c8921ba

# Data Preparation & Processing

 * FSMTaken stops being recorded in 2018
 * Schools are recorded multiple times with different names
 * 9999 is used as a value for FSMTaken and TotalFSM when count <5 upto 2017

# Analysis

Load core libraries

```{r}
library(tidyverse)  #helps wrangle data
library(ggplot2)  #helps visualize data
```

Load data:

```{r}
fsmraw <- read_csv("csv/fsmdetails.csv")
```

Remove FSMTaken as it stops being recorded in 2018:

```{r}
fsmdetails <- fsmraw %>% select(-FSMTaken)
```

Create pivot table 

```{r}
fsmdetails %>%
  group_by(SchoolName, Year) %>%
  summarise(SchoolName)

```

Wrong = Right
All Saints RC = All Saints RC School	
Applefields = Applefields School
Archbishop Holgate's CE = Archbishop Holgates CE School	
Archbishop of York's CE Junior School = Archbishop Of Yorks Ce Junior
Badger Hill Primary = Badger Hill Primary School
Canon Lee	= Canon Lee Specialist Arts College
Fulford = Fulford School	
Hob Moor Oaks	= Hob Moor Oaks School
Huntington = Huntington Primary School	
Huntington Primary School	 = Huntington Primary Academy	
Joseph Rowntree	= Joseph Rowntree School	
Lord Deramore's VC Primary School	= Lord Deramores Primary School	
Manor CE = Manor CE Academy		
Millthorpe = Millthorpe School	
Our Lady Queen of Martyrs Primary School = Our Lady Queen of Martyrs RC Primary School		
Park Grove Primary School	= Park Grove Primary Academy	
Robert Wilkinson Academy Primary School	= Robert Wilkinson Primary School	
Robert Wilkinson Primary Academy = Robert Wilkinson Primary School
St Aelred's RCVA Primary = St Aelred's RC Primary School	
St Aelreds Rc Primary School = St Aelred's RC Primary School	
St Aelreds RC Primary School = St Aelred's RC Primary School
St Barnabas C Of E Primary School = St Barnabas' CE Primary School	
St Barnabas CE Primary School	= St Barnabas' CE Primary School
St Georges Rc Primary School = St George's RC Primary School	
St Georges RC Primary School = St George's RC Primary School
St Lawrence's Primary School = St Lawrence's CE Primary School	
St Lawrences CE Primary School = St Lawrence's CE Primary School
St Marys CE Primary School = St Mary's CE Primary School
St Marys Ce Primary School = St Mary's CE Primary School
St Oswald's Primary School = St Oswald's CE Primary School	
St Oswalds Ce Primary School = St Oswald's CE Primary School	
St Oswalds CE Primary School = St Oswald's CE Primary School
St Pauls Ce Primary School = St Paul's CE Primary School	
St Pauls CE Primary School = St Paul's CE Primary School
St Wilfrid's Primary School	= St Wilfrid's RC Primary School
Stockton-on-the-Forest Primary School	= Stockton On The Forest Primary School	
Vale of York = Vale of York Academy		
Westfield Primary Community School = Westfield Community Primary School
Wheldrake CE (Aided) School	= Wheldrake with Thorganby CE Primary School	
Wheldrake with Thorganby C.E. (Aided) School = Wheldrake with Thorganby CE Primary School	
York High	= York High School	

# Key findings

# Supporting visualisations

# Top 3 Recommendations